---
title: "Mini-Project 02"
output: 
  html_document:
    keep_md: true
    toc: true
    toc_float: true
---

# Data Visualization Project 02

# Importing Data Set

Here I am importing the _spacex.csv_ data set into the markdown file, and initializing my libraries. This data set looks at launch data taken from 2010-0218. It contains information including date, launch site, payload size, mission outcome and more.
```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(sf)
library(plotly)
library(dplyr)

spacex <- read_csv("../data/spacex.csv") %>%
 mutate(
    Date       = as.Date(Date, "%B %d, %Y"),         # e.g. "June 4, 2010"
    Year       = year(Date),                         # extract integer year
    success    = (`Mission Outcome` == "Success"),   # TRUE/FALSE
    payload_kg = parse_number(`Payload Mass (kg)`)   # strip commas, make numeric
  )
glimpse(spacex)
```

## Data Summary

In this section I filtered the SpaceX data to see:
- Number of launches per year
- Success rate per launch site
- Average payload mass by booster version

```{r, message=FALSE, warning=FALSE}
#Launches per year
launches <- spacex %>%
  group_by(
    Year = as.integer(format(as.Date(Date, "%B %d, %Y"),"%Y"))
  ) %>%
  summarize(
    `Number of Launches` = n(),
    `Success Rate`       = mean(success, na.rm = TRUE)
  ) %>%
  arrange(Year)

launches


#Success rate
success <- spacex %>%
  group_by(`Launch Site`) %>%
  summarize(
    `Total Launches` = n(),
    `Success Rate`   = mean(`Mission Outcome` == "Success", na.rm = TRUE),
    .groups = "drop"
  ) %>%
  arrange(desc(`Total Launches`))

success


#Avg Mass
payload <- spacex %>%
  group_by(Booster = `Booster Version`) %>%
  summarise(
    `Average Payload (kg)` = mean(parse_number(`Payload Mass (kg)`), na.rm = TRUE),
    `Number of Launches`   = n(),
    .groups = "drop"
  ) %>%
  arrange(desc(`Average Payload (kg)`)) %>%
  filter(!is.na(`Average Payload (kg)`))
  
payload  
```

## Interactive Data Visualization

```{r, message=FALSE, warning=FALSE}
spacex <- spacex %>% 
  mutate(payload_kg = parse_number(`Payload Mass (kg)`))

plot_ly(
  data = spacex,
  x    = ~as.numeric(`Flight Number`),
  y    = ~payload_kg,
  size = ~payload_kg,
  color   = ~`Launch Site`,
  text = ~paste0(
    "Flight #: ", `Flight Number`, "<br>",
    "Date: ", Date, "<br>",
    "Booster: ", `Booster Version`, "<br>",
    "Site: ", `Launch Site`, "<br>",
    "Payload: ", payload_kg, " kg<br>",
    "Outcome: ", `Mission Outcome`
  ),
  hoverinfo = "text",
  mode = "markers"
) %>%
  layout(
    title = "SpaceX Payload by Flight",
    xaxis = list(title = "Flight Number"),
    yaxis = list(title = "Payload Mass (kg)"),
    legend = list(title = list(text = "<b>Launch Site</b>"))
  )
```

## Spatial Visualization

```{r, message=FALSE, warning=FALSE}

world_shapes <- read_sf("../data/ne_110m_admin_0_countries/ne_110m_admin_0_countries.shp", quiet = TRUE) %>% 
  filter(ISO_A3 == "USA")

#site coords
site_coords <- tribble(
  ~site,           ~lon,        ~lat,
  "CCAFS LC-40",  -80.577366,  28.561871,
  "CCAFS SLC-40", -80.583333,  28.583330,
  "KSC LC-39A",   -80.604333,  28.608389,
  "VAFB SLC-4E", -120.610829,  34.632092
)


success_by_site <- spacex %>%
  group_by(site = `Launch Site`) %>%
  summarise(
    total_launches = n(),
    success_rate   = mean(`Mission Outcome` == "Success", na.rm = TRUE),
    .groups = "drop"
  )

sites_sf <- success_by_site %>%
  left_join(site_coords, by = "site") %>%
  filter(!is.na(lon), !is.na(lat)) %>%
  mutate(
    lon = lon + runif(n(), -0.1, 0.1),
    lat = lat + runif(n(), -0.1, 0.1)
  ) %>%
  st_as_sf(coords = c("lon", "lat"), crs = 4326)

#Plot
ggplot() +
  geom_sf(data = world_shapes, fill = "gray95", color = "gray80") +
  geom_sf(
    data = sites_sf,
    aes(size = total_launches, color = success_rate),
    alpha = 0.8
  ) +
  scale_size(range = c(1, 5)) +
  scale_color_viridis_c(labels = scales::percent_format(accuracy = 1)) +
  labs(
    title    = "SpaceX Launch Sites in the USA (2010–2021)",
    subtitle = "Point size = total launches, color = success rate",
    size     = "Total\nLaunches",
    color    = "Success\nRate"
  ) +
  theme_minimal(base_size = 14) +
  theme(
    panel.grid.major = element_line(color = "white"),
    plot.title       = element_text(face = "bold", size = 16),
    plot.subtitle    = element_text(size = 12)
  )

```

## Visualization of a Model

```{r, message=FALSE, warning=FALSE}
library(broom)       # for tidy()
library(dotwhisker)  # for dwplot()


# avoid lubridate::year() name clash
spacex2 <- spacex %>%
  rename(launch_year = Year)

# fit the logistic model
log_mod <- glm(
  success ~ payload_kg + launch_year,
  data   = spacex2,
  family = binomial
)

ggplot(spacex, aes(x = payload_kg, y = as.numeric(success))) +
  geom_jitter(
  aes(x     = payload_kg,
      y     = as.numeric(success),
      colour = success),   # <-- map colour to the success flag
  height = 0.02,
  alpha  = 0.6,
  size   = 2
) +
scale_colour_manual(
  "Launch Outcome",
  values = c(
    "TRUE"  = "forestgreen",
    "FALSE" = "firebrick"
  )
) +
  stat_smooth(
  method      = "glm",
  method.args = list(family = "binomial"),
  se          = TRUE,
  color       = "forestgreen",
  fill        = alpha("palegreen", 0.3)
) +
  scale_y_continuous(labels = scales::percent_format()) +
  labs(title = "Success Probability vs. Payload Mass",
       x = "Payload Mass (kg)",
       y = "Probability of Success") +
  theme_minimal()

ggsave("figures/success_prob_vmass.png",
       width = 6, height = 4, units = "in", dpi = 300)

tidy(log_mod, conf.int = TRUE) %>%
  filter(term != "(Intercept)") %>%
  dwplot(point_size = 3) +
  labs(title = "Predictors of Launch Success",
       x = "Log-Odds Coefficient",
       y = "") +
  theme_minimal()
```

The first visualization (“Success Probability vs. Payload Mass”) shows a fitted logistic curve overlaid on the raw jittered success/failure points. Notice how the blue line sits up near 100 % success for almost the entire range of payloads and only very slightly dips at the highest masses. In other words, payload mass really doesn’t effect whether a flight succeeds or not (the 95 % confidence ribbon is very wide at the far end but the mean curve is essentially flat).

The second visualization (“Predictors of Launch Success”) is a classic coefficient‐plot of your glm. It shows the log-odds estimates (with confidence intervals) for both payload_kg and launch_year, excluding the intercept. Both coefficients sit close to zero and their intervals overlap zero, indicating neither payload nor year is a strong predictor of success.